Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-12-15 paper/main.tex
fixed typos

In light of the ongoing NIST competition for selecting the new, standard
cryptographic hash function named SHA-3, 14 of the second round candidates were
tried out in the Tahoe-LAFS file system using automated tests to measure the
performance differences. Only the immutable file support was tested.

@@

In general, four of the candidates perform better than the currently used
SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd.  By utilizing the
performance improvements given by SHA-3, Tahoe-LAFS could currently save $5\%$
on upload operations by changing hashing function to the best performing
candidate.

@@

\begin{acronym}
\acro{AES}{Advanced Encryption Standard}
\acro{API}{Application programming interface}
\acro{BMW}{Blue Midnight Wish}
\acro{CEB}{Capability Extension Block}
\acro{CPU}{Central Processing Unit}
\acro{Distutils}{Python Distribution Utilities}
\acro{FEC}{Forward Error Correction}
\acro{GCC}{GNU Compiler Collection}
\acro{KAT}{Known Answer Test}
\acro{LAFS}{Least-Authority Filesystem}
\acro{NIST}{National Institute of Standards and Technology}
\acro{RAM}{Random Access Memory}
\acro{RAID}{Redundant Array of Independent Disks}
\acro{SHA}{Secure Hash Algorithm}
\acro{SSE2}{Streaming SIMD Extensions 2}
\acro{SSSE3}{Supplemental Streaming SIMD Extensions 3}
\acro{SUPERCOP}{System for Unified Performance Evaluation Related to
Cryptographic Operations and Primitives}
\acro{SVN}{Subversion}
\acro{UEB}{URI Extension Block}
\acro{URI}{Uniform Resource Identifier}
\end{acronym}

@@

The process of selecting what is to become the new standard for a secure hashing
algorithm is carried out by the {NIST}, and is at its final stages at the time
of writing. Fourteen candidates came through to the second round
\cite{s_nistround2}.

@@

In this project assignment, we integrate the 14 \ac{SHA}-3 candidates in
the Tahoe-\ac{LAFS} and measure the performance impact. To be able to do this,
we create Python bindings for the best suited C implementations of the SHA-3
candidates, and alter the hash function calls in the source code of Tahoe.

@@

\subsection{NIST SHA-3 Competition}
The \ac{SHA} version 3 is a coming standard set to supersede the current
standards that the \ac{SHA}-1 and the \ac{SHA}-2 family has become. The hash
function that will be known as \ac{SHA}-3 will be decided by the \ac{NIST} and
chosen between the submitted contestants to the \ac{NIST} hash competition.
At the time of writing, the current status of the competition is officially
called Round 2, with 14 of 64 candidates having ``survived'' Round 1
\cite{s_fedreg}.

>>>>

In light of the ongoing NIST competition for selecting the new, standard
cryptographic hash function named SHA-3, 14 of the second round candidates were
tried out in the Tahoe-LAFS file system using automated tests to measure the
differences in performance. Only immutable file support was tested.

@@

In general, four of the candidates perform better than the currently used
SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd.  By utilizing the
performance improvements given by SHA-3, we show that Tahoe-LAFS could
currently save up to $12,36\%$ on upload operations by changing the best 
performing candidate.

@@

\begin{acronym}
\acro{AES}{Advanced Encryption Standard}
\acro{API}{Application programming interface}
\acro{BMW}{Blue Midnight Wish}
\acro{CEB}{Capability Extension Block}
\acro{CPU}{Central Processing Unit}
\acro{Distutils}{Python Distribution Utilities}
\acro{FEC}{Forward Error Correction}
\acro{GCC}{GNU Compiler Collection}
\acro{KAT}{Known Answer Test}
\acro{LAFS}{Least-Authority Filesystem}
\acro{NIST}{National Institute of Standards and Technology}
\acro{RAID}{Redundant Array of Independent Disks}
\acro{RAM}{Random Access Memory}
\acro{SHA}{Secure Hash Algorithm}
\acro{SSE2}{Streaming SIMD Extensions 2}
\acro{SSSE3}{Supplemental Streaming SIMD Extensions 3}
\acro{SUPERCOP}{System for Unified Performance Evaluation Related to
Cryptographic Operations and Primitives}
\acro{SVN}{Subversion}
\acro{UEB}{\ac{URI} Extension Block}
\acro{URI}{Uniform Resource Identifier}
\end{acronym}

@@

The process of selecting what is to become the new standard for a secure hashing
algorithm is carried out by the {NIST}, and is in its final stages at the time
of writing. Fourteen candidates came through to the second round
\cite{s_nistround2}.

@@

In this project assignment, we integrate the 14 second round \ac{SHA}-3
candidates in the Tahoe-\ac{LAFS} and measure the performance impact. To be
able to do this, we create Python bindings for the best suited C
implementations of the SHA-3 candidates, and alter the hash function calls in
the source code of Tahoe.

@@

\subsection{NIST SHA-3 Competition}

The \ac{SHA} version 3 is a coming standard set to supersede the current
standards that the \ac{SHA}-1 and the \ac{SHA}-2 family has become. The hash
function that will be known as \ac{SHA}-3 will be decided by the \ac{NIST} and
chosen between the submitted contestants to the \ac{NIST} hash competition.  At
the of writing, the current status of the competition is in it's final round
(or round 3), with only five remaining candidates. This result was not revealed
until a week before the submission of this paper, so we are focusing on all
round 2 candidates, which there are 14 of.  \cite{s_fedreg}.
